Detection and replication of epistasis influencing transcription in humans

### 2013-07-03 docs/manuscript/nature_letter.tex

\author[1,2,*]{Gibran Hemani}
\author[1,2]{Konstantin Shakhbazov}
\author[3]{Harm-Jan Westra}
\author[4]{Anjali K Henders}
\author[1,2]{Allan F. McRae}
\author[5]{Andres Metspalu}
\author[6]{Greg Gibson}
\author[4]{Nick G Martin}
\author[5,7,8]{Tonu Esko}
\author[3]{Lude Franke}
\author[4]{Grant W Montgomery}
\author[1,2]{Peter M Visscher}
\author[1,2]{Joseph E Powell}

\affil[1]{University of Queensland Diamantina Institute, University of Queensland, Princess Alexandra Hospital, Brisbane, Queensland, Australia}
\affil[2]{The University of Queensland, Queensland Brain Institute, Brisbane, QLD, Australia}
\affil[3]{Department of Genetics, University Medical Center Groningen, University of Groningen, Hanzeplein 1, Groningen, the Netherlands}
\affil[4]{Queensland Institute of Medical Research, Brisbane, Queensland, Australia}
\affil[5]{Estonian Genome Center, University of Tartu, Tartu, 51010, Estonia}
\affil[6]{School of Biology and Centre for Integrative Genomics, Georgia Institute of Technology, Atlanta, Georgia United States of America}
\affil[7]{Divisions of Endocrinology, Children's Hospital, Boston, MA, 02115, US}
\affil[8]{Medical and Population Genetics, Broad Institute, Cambridge, MA, 02142, US}
\affil[*]{Corresponding author: g.hemani@uq.edu.au}

@@

\section{Main text}

@@

Methods are now available to overcome the computational problems involved in searching for epistasis, but its detection still remains problematic due to reduced statistical power. For example increased dependence on linkage disequilibrium (LD) between causal SNPs and observed SNPs \cite{Weir2008, Hemani2013}, increased model complexity in fitting interaction terms \cite{Marchini2005}, and more extreme significance thresholds to account for increased multiple testing \cite{Cordell2009} all make it more difficult to detect epistasis in comparison to additive effects. Thus, when combined with small genetic effect sizes, as is expected in most complex traits of interest \cite{Visscher2012}, the power to detect epistasis diminishes rapidly. There are two simple ways to overcome this problem. One is by using extremely large sample sizes \cite{LangoAllen2010}; another is by analysing traits that are likely to have large effect sizes among common variants. Because our focus was to ascertain the extent to which instances of epistasis occur amongst natural genetic variation we designed a study around the latter approach and searched for epistatic genetic effects that influence gene expression levels. Transcription levels can be measured for thousands of genes. These traits are largely heritable but on average less polygenic than high level phenotypes \cite{Powell2013}, thus it is expected that many genetic effects will be relatively large, maximising the chance at detecting epistasis, should it exist.

@@

In addition, using the meta analysis from the replication samples only, we observed that 316 of the remaining 404 discovery SNPs had replication interaction $p$-values more extreme than the 2.5\% confidence interval of the distribution under the null distribution of no epistatic effects ($p << 1.0 \times 10^{-16}$, Figure \ref{fig:qqMeta} and Supplementary Figure \ref{fig:qqMetaNonsig}). The congruence of the epistatic networks in discovery and replication datasets is shown in Figure \ref{fig:fireworks}, demonstrating that these complex genetic patterns are common even across independent datasets. A further replication was attempted using the Centre for Health Discovery and Wellbeing (CHDWB) dataset \cite{Preininger2013}, but only 27 of the SNP pairs passed filtering because the sample size was small ($n=139$), and likely due to insufficient power we found no evidence for replication. It should be noted that although it is a necessary step to establish the veracity of the signals from the discovery set, replication of epistasis is theoretically difficult because the dependence on LD between observed SNPs and causal variants is up to three orders of magnitude higher than it is for independent additive effects \cite{Weir2008, Hemani2013}. Therefore these results are encouraging with regards to the detection and replication of epistasis.

@@

Of the discovery interactions, 47 were \emph{cis}-\emph{cis} acting (both SNPs were on the same chromosome as the expression gene), 441 were \emph{cis}-\emph{trans}-acting, and 13 were \emph{trans}-\emph{trans}-acting. We observed a wide range of significant GP maps (Figure \ref{fig:gpmaps}) but the most common pattern of epistasis that we detected involved a \emph{trans}-SNP masking the effect of an additive \emph{cis}-SNP. For example, MBNL1 (involved in RNA modification and regulation of splicing \cite{Ho2004}) has a \emph{cis} effect at rs13069559 which in turn is controlled by 13 \emph{trans}-SNPs and one \emph{cis}-SNP that each exhibit a masking pattern, such that when the \emph{trans}-SNP is homozygous for the masking allele the decreasing allele of the \emph{cis}-SNP no longer has an effect (Supplementary Figure \ref{fig:MBNL1_3D}. Each of these interactions have evidence for replication in at least one dataset and six are significant at the Bonferroni level (Supplementary Figure \ref{fig:circleplots}). We see similar epistatic networks involving multiple (eight or more) \emph{trans}-acting SNPs for other gene expresson levels too, for example TMEM149 (Supplementary Figure \ref{fig:TMEM149_3D}), NAPRT1 (Supplementary Figure \ref{fig:NAPRT1_3D}), TRAPPC5 (Supplementary Figure \ref{fig:TRAPPC5_3D}), and CAST (Supplementary Figure \ref{fig:CAST_3D}). We observed that these five gene expression phenotypes had whole-genome non-additive variance component estimates within the 95th percentile of the 17,994 gene expression phenotypes that were analysed in Powell \emph{et al} 2013 (Supplementary Table \ref{tab:vc}, Methods) \cite{Powell2013}.

% \subsection{Functional annotation}
In total the 501 interactions comprised 781 unique SNPs, which we analysed for functional enrichment (Methods). We tested the SNPs for cell-type specific overlap with transcriptionally active chromatin regions, tagged by histone-3-lysine-4,3-methylation (H3K4me3) chromatin marks, in 34 cell types \cite{Trynka2013} (Supplementary Figure \ref{fig:h3k4me3}). There was significant enrichment for \emph{cis}-acting SNPs in haematopoietic cell types only ($p < 1 \times 10^{-4}$ for the three tissues with the strongest enrichment after adjusting for multiple testing). However \emph{trans}-acting SNPs did not show any tissue specific enrichment ($p > 0.1$ for all tissues). This difference between \emph{cis} and \emph{trans} SNPs suggests different roles in which epistasis might arise where the \emph{cis}-SNPs provide tissue specificity in these interactions. There is also strong enrichment for SNPs to be localised in enhancer regions \cite{Ward2012a}, consistent for both \emph{cis} and \emph{trans} SNPs ($p < 1 \times 10^{-6}$).

@@

\subsection{Methods Summary}
We searched for pairwise epistasis exhaustively in the BSGS discovery dataset \cite{Powell2012}, which comprises 846 individuals who are genotyped at 528,509 autosomal SNPs and who have gene expression levels measured in peripheral blood samples for 7,339 probes representing 6,158 RefSeq genes. Recent hardware and software \cite{Hemani2011} advances made it possible to perform the $1.03 \times 10^{15}$ statistical tests to complete this analysis. We used permutation analysis \cite{Churchill1994a} to calculate an experiment-wide significance threshold of $T_{e} = 2.91 \times 10^{-16}$ at the 5\% family-wise error rate (FWER). SNP pairs were modelled for full genetic effects, including marginal additive and dominance at both SNPs plus four interaction terms. Though we could have used a less complex model to improve statistical efficiency, we deemed it important to be agnostic about the type of epistasis that might exist, and therefore chose not to over-parameterise the test \cite{Marchini2005, Hemani2013}. Because there are many large marginal effects present in these data it was necessary to perform several filtering steps to exclude SNP pairs that were significant due to marginal effects alone. All SNP pairs with LD $r^2 > 0.1$ and $D'^{2} > 0.1$ were removed to minimise the possibility of haplotype effects. All SNP pairs were required to have at least five data points in all nine genotype classes. If multiple SNP pairs were present on the same chromosomes for a particular expression trait then only the sentinal SNP pair was retained. Finally, a nested test contrasting the full genetic model against the marginal additive and dominance model was performed for each remaining SNP pair (Methods), resulting in 501 significant interactions after Bonferroni correction for multiple testing of the filtered SNPs. The significant SNP pairs were carried forward for replication in two independent datasets that used the same expression assays for analysing transcription in peripheral blood, the Fehrmann dataset \cite{Fehrmann2011} ($n=1240$) and the Estonian Genome Centre University of the University of Tartu (EGCUT) dataset \cite{Metspalu2004} ($n=891$). Of these, 434 passed filtering in both replication datasets. A meta analysis on the interaction $p$-values from each replication dataset was performed to provide an overall replication statistic for each putative interaction.

\subsection{Acknowledgements}

@@

The University of Queensland group is supported by the Australian Research Council (DP130102666, DE130100691), the Australian National Health and Medical Research Council (APP1011506, APP1047956, APP1048853) and the National Institutes of Health (GM099568, GM075091).

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=2.5in]{gpbonfrep.pdf}
	\caption{\textbf{Replication of GP maps in two independent populations} The GP maps for each epistatic interaction that is significant at the Bonferroni level in both replication datasets are shown. Each GP map consists of nine tiles where each tile represents the expression level for that two-locus genotype class. Phenotypes are for gene transcript levels (dark coloured tiles = low expression, light coloured tiles = high expression). Columns of GP maps are for each independent population. Rows of GP maps are for each of 30 significantly replicated interactions at the Bonferroni level, corresponding to the rows in Table \ref{tab:bonferroni}. There is a clear trend of the GP maps replicating across all three datasets.}
	\label{fig:gpmaps}
\end{figure}
\clearpage

@@

\begin{figure}
	\centering
	\includegraphics[width=5in]{ChromatinStatesOverlap}
	\caption{\textbf{Location of SNPs relative to genomic features} All SNPs within 1Mb and $r^2 > 0.8$ of each \emph{cis}- and \emph{trans}-SNP were taken to find which genomic features (\emph{x}-axis) were covered by the SNPs that compose the 501 significant interactions. Green bars represent the proportion (\emph{y}-axis) of the 528,509 SNPs used in the analysis that fall within the range of the different genomic features. There is enrichment for \emph{cis}-acting SNPs (red bars) in promotor regions, but \emph{trans}-acting SNPs (blue bars) are not enriched for genomic features. The labels on the $x$-axis are as follows: TSS = Predicted promoter region including TSS, PF = Predicted promoter flanking region, E = Predicted enhancer, WE = Predicted weak enhancer or open chromatin cis regulatory element, CTCF = CTCF enriched element, T = Predicted transcribed region, R = Predicted Repressed or Low Activity region}
	\label{fig:snplocations}
\end{figure}
\clearpage

\begin{figure}
	\centering
	\includegraphics[width=5in]{501_pairs_snp_list_Grid}
	\caption{\textbf{Tissue specific enrichment of SNPs in transcriptionally active regions} The locations of transcriptional activity can be predicted by chromatin marks, assayed by H3K4me3 \cite{Trynka2013}. Enrichment $p$-values are calculated using permutation analysis for 34 different cell types ($y$-axis) in four tissue types (Rows of boxes). There is enrichment for \emph{cis}-acting SNPs in Haematopoietic tissue types only. \emph{Trans}-acting SNPs have no tissue specificity.}
	\label{fig:h3k4me3}
\end{figure}
\clearpage

@@

ABCA7 & ILMN\_1743205 & 19 & rs3752237 & 19 & 1047161 & ABCA7 & rs596183 & 6 & 158100199 &  & 5.82 & 0.38 & 0.02 & 0.09 &  \\

@@

\begin{tablenotes}
\item[a] Phenotypes are expression levels of RefSeq Genes
\item[b] Illumina probe ID used to measure gene expression
\item[c] Physical SNP position in base pairs (HG19)
\item[d] RefSeq Gene ID of gene expression level that is influenced by the SNP (BSGS discovery dataset, significance threshold $= 1.29 \times 10^{-11}$)
\item[e] Interaction $-\log_{10} p$-value from discovery dataset
\item[f] Interaction $-\log_{10} p$-value from replication dataset
\item[g] Interaction $-\log_{10} p$-value from meta analysis of replication datasets only
\item[h] Distance in Mb between interacting SNPs for \emph{cis}-\emph{cis} acting SNP pairs

@@

\begin{table}[ht]
\begin{threeparttable}
\caption{\textbf{Estimation of additive and non-additive variance components from pedigree information}}
\label{tab:vc}
\begin{tabular}{llrrrr}

>>>>

\author[1,2,*]{Gibran Hemani}
\author[1,2]{Konstantin Shakhbazov}
\author[3]{Harm-Jan Westra}
\author[4,5,6]{Tonu Esko}
\author[7]{Anjali K Henders}
\author[1,2]{Allan F. McRae}
\author[2]{Jian Yang}
\author[8]{Greg Gibson}
\author[7]{Nick G Martin}
\author[4]{Andres Metspalu}
\author[3]{Lude Franke}
\author[7,+]{Grant W Montgomery}
\author[1,2,+]{Peter M Visscher}
\author[1,2,+]{Joseph E Powell}

\affil[1]{University of Queensland Diamantina Institute, University of Queensland, Princess Alexandra Hospital, Brisbane, Queensland, Australia}
\affil[2]{The University of Queensland, Queensland Brain Institute, Brisbane, QLD, Australia}
\affil[3]{Department of Genetics, University Medical Center Groningen, University of Groningen, Hanzeplein 1, Groningen, the Netherlands}
\affil[4]{Estonian Genome Center, University of Tartu, Tartu, 51010, Estonia}
\affil[5]{School of Biology and Centre for Integrative Genomics, Georgia Institute of Technology, Atlanta, Georgia United States of America}
\affil[6]{Divisions of Endocrinology, Children's Hospital, Boston, MA, 02115, US}
\affil[7]{Queensland Institute of Medical Research, Brisbane, Queensland, Australia}
\affil[8]{Medical and Population Genetics, Broad Institute, Cambridge, MA, 02142, US}
\affil[+]{These authors contributed equally}
\affil[*]{Corresponding author: g.hemani@uq.edu.au}

@@

\section*{Main text}

@@

Methods are now available to overcome the computational problems involved in searching for epistasis, but its detection still remains problematic due to reduced statistical power. For example increased dependence on linkage disequilibrium (LD) between causal SNPs and observed SNPs \cite{Weir2008, Hemani2013}, increased model complexity in fitting interaction terms \cite{Marchini2005}, and more extreme significance thresholds to account for increased multiple testing \cite{Cordell2009} all make it more difficult to detect epistasis in comparison to additive effects. Thus, when combined with small genetic effect sizes, as is expected in most complex traits of interest \cite{Visscher2012}, the power to detect epistasis diminishes rapidly. There are two simple ways to overcome this problem. One is by using extremely large sample sizes \cite{LangoAllen2010}; another is by analysing traits that are likely to have large effect sizes among common variants. Because our focus was to ascertain the extent to which instances of epistasis occur amongst natural genetic variation we designed a study around the latter approach and searched for epistatic genetic effects that influence gene expression levels. Transcription levels can be measured for thousands of genes. These traits are largely heritable but on average less polygenic than high level phenotypes \cite{Powell2013}, thus many genetic effects are relatively large, maximising the chance at detecting epistasis, should it exist.

@@

In addition, using the meta analysis from the replication samples only, we observed that 316 of the remaining 404 discovery SNPs had replication interaction $p$-values more extreme than the 2.5\% confidence interval of the distribution under the null distribution of no epistatic effects ($p << 1.0 \times 10^{-16}$, Figure \ref{fig:qqMeta} and Supplementary Figure \ref{fig:qqMetaNonsig}). The congruence of the epistatic networks in discovery and replication datasets is shown in Figure \ref{fig:fireworks}, demonstrating that these complex genetic patterns are common even across independent datasets. A further replication was attempted using the Centre for Health Discovery and Wellbeing (CHDWB) dataset \cite{Preininger2013}, but only 27 of the SNP pairs passed filtering because the sample size was small ($n=139$), and likely due to insufficient power we found no evidence for replication. It should be noted that although it is a necessary step to establish the veracity of the signals from the discovery set, replication of epistasis is difficult in practice because the dependence on LD between observed SNPs and causal variants is up to three orders of magnitude higher than it is for independent additive effects \cite{Weir2008, Hemani2013}. Therefore these results are encouraging with regards to the detection and replication of epistasis.

@@

Of the discovery interactions, 47 were \emph{cis}-\emph{cis} acting (both SNPs were on the same chromosome as the expression gene), 441 were \emph{cis}-\emph{trans}-acting, and 13 were \emph{trans}-\emph{trans}-acting. We observed a wide range of significant GP maps (Figure \ref{fig:gpmaps}) but the most common pattern of epistasis that we detected involved a \emph{trans}-SNP masking the effect of an additive \emph{cis}-SNP. For example, MBNL1 (involved in RNA modification and regulation of splicing \cite{Ho2004}) has a \emph{cis} effect at rs13069559 which in turn is controlled by 13 \emph{trans}-SNPs and one \emph{cis}-SNP that each exhibit a masking pattern, such that when the \emph{trans}-SNP is homozygous for the masking allele the decreasing allele of the \emph{cis}-SNP no longer has an effect (Supplementary Figure \ref{fig:MBNL1_3D}. Each of these interactions have evidence for replication in at least one dataset and six are significant at the Bonferroni level (Supplementary Figure \ref{fig:circleplots}). We see similar epistatic networks involving multiple (eight or more) \emph{trans}-acting SNPs for other gene expresson levels too, for example TMEM149 (Supplementary Figure \ref{fig:TMEM149_3D}), NAPRT1 (Supplementary Figure \ref{fig:NAPRT1_3D}), TRAPPC5 (Supplementary Figure \ref{fig:TRAPPC5_3D}), and CAST (Supplementary Figure \ref{fig:CAST_3D}). We observed that from pedigree analysis these five gene expression phenotypes had non-additive variance component estimates within the 95th percentile of the 17,994 gene expression phenotypes that were analysed previously \cite{Powell2013} (Supplementary Table \ref{tab:vc}, Methods).

% \subsection{Functional annotation}
In total the 501 interactions comprised 781 unique SNPs, which we analysed for functional enrichment (Methods). We tested the SNPs for cell-type specific overlap with transcriptionally active chromatin regions, tagged by histone-3-lysine-4,3-methylation (H3K4me3) chromatin marks, in 34 cell types \cite{Trynka2013} (Supplementary Figure \ref{fig:h3k4me3}). There was significant enrichment for \emph{cis}-acting SNPs in haematopoietic cell types only ($p < 1 \times 10^{-4}$ for the three tissues with the strongest enrichment after adjusting for multiple testing). However \emph{trans}-acting SNPs did not show any tissue specific enrichment ($p > 0.1$ for all tissues). This difference between \emph{cis} and \emph{trans} SNPs suggests different roles in epistatic interactions where the \emph{cis}-SNPs provide tissue specificity in these interactions. There is also strong enrichment for SNPs to be localised in enhancer regions \cite{Ward2012a}, consistent for both \emph{cis} and \emph{trans} SNPs ($p < 1 \times 10^{-6}$).

@@

\subsection*{Methods Summary}
We searched for pairwise epistasis exhaustively in the BSGS discovery dataset \cite{Powell2012}, which comprises 846 individuals who are genotyped at 528,509 autosomal SNPs and who have gene expression levels measured in peripheral blood samples for 7,339 probes representing 6,158 RefSeq genes. Recent hardware and software \cite{Hemani2011} advances that use graphics processing units (GPUs) made it possible to perform the $1.03 \times 10^{15}$ statistical tests to complete this analysis. We used permutation analysis \cite{Churchill1994a} to calculate an experiment-wide significance threshold of $T_{e} = 2.91 \times 10^{-16}$ at the 5\% family-wise error rate (FWER). SNP pairs were modelled for full genetic effects, including marginal additive and dominance at both SNPs plus four interaction terms. Though we could have used a less complex model to improve statistical efficiency, we deemed it important to be agnostic about the type of epistasis that might exist, and therefore chose not to over-parameterise the test \cite{Marchini2005, Hemani2013}. Because there are many large marginal effects present in these data it was necessary to perform several filtering steps to exclude SNP pairs that were significant due to marginal effects alone. All SNP pairs with LD $r^2 > 0.1$ and $D'^{2} > 0.1$ were removed to minimise the possibility of haplotype effects. All SNP pairs were required to have at least five data points in all nine genotype classes. If multiple SNP pairs were present on the same chromosomes for a particular expression trait then only the sentinal SNP pair was retained. Finally, a nested test contrasting the full genetic model against the marginal additive and dominance model was performed for each remaining SNP pair (Methods), resulting in 501 significant interactions after Bonferroni correction for multiple testing of the filtered SNPs. The significant SNP pairs were carried forward for replication in two independent datasets that used the same expression assays for analysing transcription in peripheral blood, the Fehrmann dataset \cite{Fehrmann2011} ($n=1240$) and the Estonian Genome Centre University of the University of Tartu (EGCUT) dataset \cite{Metspalu2004} ($n=891$). Of these, 434 passed filtering in both replication datasets. A meta analysis on the interaction $p$-values from each replication dataset was performed to provide an overall replication statistic for each putative interaction.

\subsection*{Acknowledgements}

@@

The University of Queensland group is supported by the Australian National Health and Medical Research Council (NHMRC) grants 389892, 496667, 613601, 1010374 and 1046880, the Australian Research Council (ARC) grant (DE130100691), and by National Institutes of Health (NIH) grants GM057091 and GM099568.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=2.5in]{gpbonfrep.pdf}
	\caption{\textbf{Replication of GP maps in two independent populations} The GP maps for each epistatic interaction that is significant at the Bonferroni level in both replication datasets are shown. Each GP map consists of nine tiles where each tile represents the expression level for that two-locus genotype class. Phenotypes are for gene transcript levels (dark coloured tiles = high expression, light coloured tiles = low expression). Columns of GP maps are for each independent population. Rows of GP maps are for each of 30 significantly replicated interactions at the Bonferroni level, corresponding to the rows in Table \ref{tab:bonferroni}. There is a clear trend of the GP maps replicating across all three datasets.}
	\label{fig:gpmaps}
\end{figure}
\clearpage

@@

\begin{figure}
	\centering
	\includegraphics[width=5in]{ChromatinStatesOverlapFixedOrder}
	\caption{\textbf{Location of SNPs relative to genomic features} All SNPs within 1Mb and $r^2 > 0.8$ of each \emph{cis}- and \emph{trans}-SNP were taken to find which genomic features (\emph{x}-axis) were covered by the SNPs that compose the 501 significant interactions. Green bars represent the proportion (\emph{y}-axis) of the 528,509 SNPs used in the analysis that fall within the range of the different genomic features. There is enrichment for \emph{cis}-acting SNPs (red bars) in promotor regions, but \emph{trans}-acting SNPs (blue bars) are not enriched for genomic features. The labels on the $x$-axis are as follows: E = Predicted enhancer, CTCF = CTCF enriched element, WE = Predicted weak enhancer or open chromatin cis regulatory element, PF = Predicted promoter flanking region, TSS = Predicted promoter region including transcriptional start site, T = Predicted transcribed region, R = Predicted Repressed or Low Activity region}
	\label{fig:snplocations}
\end{figure}
\clearpage

\begin{figure}
	\centering
	\includegraphics[width=5in]{H3K4me3Enrichment501_list}
	\caption{\textbf{Tissue specific enrichment of SNPs in transcriptionally active regions} The locations of transcriptional activity can be predicted by chromatin marks, assayed by H3K4me3 \cite{Trynka2013}. Enrichment $p$-values are calculated using permutation analysis for 34 different cell types ($y$-axis) in four tissue types (Rows of boxes). There is enrichment for \emph{cis}-acting SNPs in Haematopoietic tissue types only. \emph{Trans}-acting SNPs have no tissue specificity.}
	\label{fig:h3k4me3}
\end{figure}
\clearpage

@@

ABCA7 & ILMN\_1743205 & 19 & rs3752237 & 19 & 1047161 & ABCA7 & rs596183 & 6 & 158100199 &  & 5.82 & 0.38\tnote{i} & 0.02\tnote{i} & 0.09\tnote{j} &  \\

@@

\begin{tablenotes}
\item[a] Phenotypes are expression levels of RefSeq Genes
\item[b] Illumina probe ID used to measure gene expression
\item[c] Physical SNP position in base pairs (HG19)
\item[d] RefSeq Gene ID of gene expression level that is influenced by the SNP (BSGS discovery dataset, significance threshold $= 1.29 \times 10^{-11}$)
\item[e] Interaction $-\log_{10} p$-value from discovery dataset
\item[f] Interaction $-\log_{10} p$-value from replication dataset
\item[g] Interaction $-\log_{10} p$-value from meta analysis of replication datasets only
\item[h] Distance in Mb between interacting SNPs for \emph{cis}-\emph{cis} acting SNP pairs
\item[i] $p$-values are absent if the interaction did not pass the QC filtering in the replication dataset
\item[j] Meta analysis $p$-values are absent if the interaction did not pass the QC filtering in either replication dataset

@@

\begin{table}[ht]
\begin{threeparttable}
\caption{\textbf{Estimation of additive and non-additive variance components from pedigree information} Taken from previous analysis in Powell et al 2013 \cite{Powell2013}}
\label{tab:vc}
\begin{tabular}{llrrrr}
